Inspiring Confidence and Easing Insecurity in the Pre-election period: The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections

### 2025-03-19 docs/_tex/index.tex
draft 3 complete. Also full project render

% ---------------------------
% END CUSTOM PREAMBLE STUFF
% ---------------------------
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{dcolumn} 
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={title},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={RoyalBlue},
  pdfcreator={LaTeX via pandoc}}

\title{title}
% subtitles do not seem to work with article class?
%%\usepackage{etoolbox}
%\makeatletter
%\providecommand{\subtitle}[1]{% add subtitle to \maketitle
%  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
%}
%\makeatother
%%\subtitle{subtitle}

@@

% add the short title to the fancy header
\fancyhead[R]{A Short title}
\fancyhead[L]{Isaiah}

@@

\setstretch{1.25}
\pagebreak

@@

Election administration officials make efforts to sustain public trust
and confidence in the fairness and accuracy of elections, and attempt to
boost such confidence where it may be deprived. Concerns for safety have
developed among election staff and voters in more recent elections.
Regular measures are taken to enhance the
\emph{trustworthiness}\footnote{I adopt a distinction made between
  \emph{trustworthiness} and \emph{trust} in elections
  (\citeproc{ref-stewart2022}{Stewart 2022}). The ``\emph{worthiness}''
  of one's trust in the conduct and administration of elections is based
  on the extent that outcomes of an election reasonably follow the rules
  prescribed and can be adjudicated as such. The public's \emph{trust}
  in elections, however, is amendable to an indefinite number of factors
  that may be unrelated to the formal structure or procedures of
  election administration. I can recognize that my car is trustworthy
  prior to ever driving it because it is structurally sound; I have
  every reason to deem it \emph{worthy of my trust}. However, I don't
  trust my car because I am sure it is haunted.} of the electoral
process through practices meant to improve the conduct, transparency, or
overall administration of elections in the United States.

Although election officials undertake great efforts to enhance the
trustworthiness of election administration, public \emph{trust} in
elections is a psychological construct influenced by many things outside
of election official control such as partisanship or elite rhetoric
(\citeproc{ref-hooghe2018}{Hooghe 2018};
\citeproc{ref-sances2015}{Sances and Stewart 2015}). Moreover, a
person's evaluation of the election in hindsight is often influenced by
the election outcome itself (\citeproc{ref-daniller2019}{Daniller and
Mutz 2019}; \citeproc{ref-stewart2022}{Stewart 2022}). Thus, measures
taken by election officials can be undermined, trivialized, or made
irrelevant depending on how one feels after the election results have
come out.

Such volatile attitudes and evaluations post-election can leave a
lasting impression that election officials must contend with upon the
next election cycle (\citeproc{ref-bowler2024}{Bowler and Donovan 2024};
\citeproc{ref-levendusky2024}{Levendusky et al. 2024}). For instance, we
have witnessed many people's outright refusal to accept the 2020 U.S.
election results as legitimate despite consistent review of the evidence
confirming the results as fair and accurate. Such a case demonstrates
that public trust in elections is, at best, only partial to
trustworthiness of election administration in the United
States\footnote{There's little that could justify a conceptual
  distinction between public trust pre-election and public trust
  post-election. The temporal element renders the difference between
  pre- and post-election more operative than conceptual at a
  foundational level.}.

@@

Since the results and events following the 2020 election loomed large in
anticipation of the U.S. 2024 general elections, it is reasonable to
take into account beliefs about legitimacy of the 2020 election. This
group who denies the legitimacy of the 2020 election results were
possibly the most apprehensive (distrusting) of the 2024 election.
However, given that veteran service members are held in high regard,
generally speaking, announcing that veterans are being actively
recruited to work as election staff and volunteers is expected to
positively influence confidence in elections among those who remained
firm in the belief that the 2020 election results were illegitimate. I
take into account that legitimacy beliefs about the 2020 election and
expectations of electoral fraud are related to partisanship and include
partisanship as a control in the analysis\footnote{To be clear, I
  conducted an auxiliary regression analysis to examine the relationship
  between legitimacy beliefs and partisanship. Results revealed a
  statistically significant relationship between the two variables and
  thus omitting either would result in biased estimates.}.

In addition, I expect that both expectations of electoral fraud and
concerns for voter safety would lessen (shrink) when presented with
information that veterans are being actively recruited to work and
volunteer in election offices.

Finally, prior research has shown that confidence that one's own vote
would be counted as intended has been much stronger than confidence that
votes nationwide would be treated likewise
(\citeproc{ref-sances2015}{Sances and Stewart 2015};
\citeproc{ref-stewart2022}{Stewart 2022}). In other words, the public is
more trusting of elections closer to home. Taking into account that
confidence is lower for elections conducted outside of one's local area,
then information that veterans are being recruited to help conduct
elections in locations beyond one's locale may close this gap between
confidence in elections within one's local area and confidence in
elections elsewhere. Although it may be difficult to discern whether
recruiting veteran service members would improve confidence in elections
near one's local area in light of such hometown favoritism, a positive
effect on confidence in elections held beyond one's local area---where
confidence is already expectedly lower---can help discern the impact of
the group alone on public opinion. Therefore, any observed disparity of
confidence in elections (i.e., the ``confidence gap'') between elections
within one's local area and elections beyond will be smaller among those
who are presented with information about election official recruitment
efforts of veteran service members.

@@

Survey participants read either a treatment or control vignette, which
resembled a news article about efforts in Maricopa County, AZ to recruit
election staff and volunteers for the 2024 general election\footnote{The
  article vignettes were written and designed by Michael Hanmer Ph.D and
  myself with input from We the Veterans, a non-profit organization}.
The treatment vignette referred to a program designed to recruit
veterans and their family members and describes an interviewee ``Jordan
Braxton'' as an Army veteran. The control vignette simply omitted any
mention of veterans and their family members, and didn't describe
``Jordan Braxton'' as an Army veteran. Beyond those small differences
and the headlines, the article vignettes are identical. Therefore,
effects can be attributed to the information about veterans in the
treatment vignette\footnote{Complete text of treatment and control
  vignettes are included in the Appendix}.

@@

\begingroup\fontsize{8}{10}\selectfont

\begin{tabu} to \linewidth {>{\raggedright\arraybackslash}p{8cm}>{\raggedright\arraybackslash}p{8cm}}
\toprule
Trust & Distrust\\
\midrule
How confident are you that votes in Maricopa County, AZ will be counted as voters intend in the elections this November? & There will be voter fraud, that is, people who are not eligible to vote will vote, or vote more than once\\
\addlinespace
How confident are you that election officials, their staff, and volunteers in Maricopa County, AZ will do a good job conducting the elections this November? & Many votes will not actually be counted\\
\addlinespace
How commited do you think election staff and volunteers in Maricopa County, AZ will be to making sure the elections are fair and accurate? & Many people will show up to vote and be told they are not eligible\\
\addlinespace
How confident are you that the voting process will be fair in Maricopa County, AZ? & A foreign country will tamper with the votes cast in this area to change the results\\
\addlinespace
How confident are you that election systems in Maricopa County, AZ will be secure from hacking and other technological threats? & Election officials in Maricopa County, Arizona will try to discourage some people from voting\\
\bottomrule
\end{tabu}
\endgroup{}

@@

Upon construction of the two summated scales of \emph{trust} and
\emph{distrust}, internal consistency reliability coefficient \(\alpha\)
were estimated for both and found to be adequate (\(\alpha\)
\textgreater{} 0.8); the trust in elections scale \(\alpha\) = 0.93 and
the distrust scale \(\alpha\) = 0.86. In other words, the items in each
of the scales were intercorrelated well enough for each to compose a
single scale presuming a unidimensional construct for each\footnote{See
  Appendix C for analysis.}.

@@

In light of this, I reverse coded items on the distrust scale and
combined responses to the trust in elections scale. That is, item
responses on the trust scale ranged from 0 (``Not confident at all'') to
3 (``Very confident), whereas responses on the distrust scale ranged
from 0 (''Not likely at all'') to -3 (``Very likely''). Combining
responses in this way resulted in an overall measure of \emph{confidence
in elections} to serve as the dependent variable\footnote{The
  methodological concerns between the use of (unit weighted) sum or mean
  scores compared to the use of estimated factor scores for Likert
  scales composed of multiple Likert items are legion. The use of either
  computed sum scores or estimated factor scores as a general practice
  is continuously debated (\citeproc{ref-mcneish2023}{McNeish 2023};
  \citeproc{ref-mcneish2020}{McNeish and Wolf 2020};
  \citeproc{ref-widaman2022}{Widaman and Revelle 2022},
  \citeproc{ref-widaman2024}{2024}), yet there is no general consensus
  on the most appropriate method generally. The use of either appears to
  largely depend on the researcher's particular research objective,
  theoretical suppositions, and necessity. Another common concern is the
  inappropriate treatment of ordinal variables as interval which levy
  strong assumptions, e.g., each item contributes equally to the
  variable being measured and that ordinal response categories are
  equidistant as though on an interval scale
  (\citeproc{ref-wang2013}{Wang et al. 2013}; for contrast, see
  \citeproc{ref-robitzsch2020}{Robitzsch 2020}).}. On this scale,
positive values reflect higher confidence in elections, negative values
reflect greater distrust, and values closer to zero correspond to an
individual's degree of \emph{insecurity}.

@@

I first review responses to the survey items capturing trust and
distrust separately by treatment condition. When distinguished by
experiment condition, trust in elections admin is noticeably higher for
those who read the treatment vignette. The difference is more pronounced
for some items over others. That being said, there is a definite
distinction between items pertaining to Maricopa County, AZ compared to
items that pertain to a respondent's local area. In fact, as expected,
the treatment appears to have hardly made a difference when the items
concerned one's local area.

\begin{figure}[h]

\caption{\label{fig-patch10}Trust in Election Administration by
Experiment Condition}

\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-figures-fig-patch10-output-1.png}}

Reviewing item responses distinguished by treatment condition reveals
particulars of the treatment effect (Figure~\ref{fig-patch10}). Among
the items that asked about Maricopa County, AZ, it is a bit easier to
see upon which items the treatment vignette had the greatest effect. It
appears that the main effect of the treatment is upon responses to the
item that measures confidence in the security of election technology.

Regarding confidence in the commitment of election staff and volunteers,
a higher proportion of respondents in the treatment group selected
``Very confident'' over ``Somewhat confident'' compared to the control
group, but only for items concerning Maricopa County, AZ. Although
confidence in election workers was generally high across the board, the
shift from ``Somewhat'' to ``Very'' confident for items pertaining to
election worker \emph{commitment} to fairness and accuracy seems to
clearly speak to the group being recruited, i.e., veterans and family
members. However, the same visible shift in proportion from ``Somewhat''
to ``Very'' confident isn't observed when items asked about one's local
area. It may be the case that respondents harbor particular attitudes or
sentiment about Maricopa County, AZ itself, or there may be nothing
uniquely special about Maricopa County, AZ supposing that elections
anywhere beyond one's local area generally garner slightly less
confidence.

What this does suggest, however, is that trust in the fairness and
accuracy of the electoral process is notably higher upon the notion that
military veterans will be directly engaged as election staff and
volunteers in places where such trust falters. The question turns to
whether this also lowers distrust, or the extent to which one expects
electoral fraud to occur.

\caption{\label{fig-patch11}Disrust in Election Administration by
Experiment Condition}

\pandocbounded{\includegraphics[keepaspectratio]{index_files/figure-latex/notebooks-figures-fig-patch11-output-1.png}}

Similar to the trust scale items, the items on the distrust scale were
associated with lower distrust in the treatment condition compared to
the control. A quick glance at Figure~\ref{fig-patch11} shows what
appears to be a significant reduction in distrust, especially regarding
the expectations that many votes wouldn't be counted, many voters would
be turned away, and that election officials would discourage voters from
casting a ballot. Interestingly, the expectation that there would be
foreign interference in elections in Maricopa County, AZ were pretty
resilient against the treatment. And again, responses for items
pertaining to elections in one's local area are practically
indistinguishable between treatment and control.

\subsubsection{Confidence in elections}\label{confidence-in-elections}

Once trust and distrust were composed into unified variable of
confidence in elections, I then examined distribution of the dependent
variable followed by analysis of the effect of the treatment by
comparison of mean differences between experiment conditions. Recall
that on this scale, zero reflects insecurity (e.g., similar or equal
values of trust and distrust), whereas values diverging from zero
reflect confidence in either direction, trust or distrust respectively.
Positive values reflect trust in elections administration (e.g.,
elections will be fair and accurate), whereas negative values reflect
distrust in elections (i.e., expectation that election fraud will
occur). To retain a meaningful zero, I rescaled the sum scores to range
from -3 to 3\footnote{This range is mostly arbitrary, as a range from -1
  to 1 works much the same. When relying on simple sum total score, the
  scale ranges from -15 to 15. So a single point increase from 0 to 1
  may reflect a combination of a single ``Very confident'' response on a
  trust item compensated by a single ``Somewhat likely'' response on a
  distrust item (i.e., 3-2 = 1), or some other equivalent combination.
  Although the use of mean scores would place scores back onto the
  response scale metric (e.g., from 0 to 3, reflective of ``not at all''
  to ``very''), this results in a unipolar scale (from 0 to 3). The
  bipolar scale resulting from the composition of positive trust and
  negative distrust engenders meaning to zero and negative values. As
  such, composite mean scores become inappropriate and interpretation of
  scores would no longer be feasible. In other words, negative values on
  the scale for confidence in elections hold substantive meaning (e.g.,
  distrust, absent trust), which makes transforming scores to fit a
  unipolar scale inappropriate.} and then superimposed two histograms of
confidence scores: one derived from AZ items and another for scores
derived from local area items. Doing this is illustrative of the
relative distribution of confidence that differed depending on the
location of elections referred to in the survey items (e.g., Maricopa
County, AZ).

As revealed by review of the item responses, the distribution of
confidence differed depending on whether survey items pertained to
elections in Maricopa County, AZ, or elections within one's local area.

@@

Overall, the sample expressed more trust in elections relative to
distrust seeing as how the distribution of confidence is largely
positive (Figure~\ref{fig-dist}). However, there's a clear difference
between confidence in Maricopa county, AZ elections and elections within
one's local area. A paired t-test confirms that this difference was
statistically significant, though somewhat small (mean difference
\(= 0.325\), \(95\%\) CI \([0.27, 0.38]\), \(t(1286) = 11.61\),
\(p < .001\)). Nonetheless, respondents appeared less confident
regarding elections in Maricopa County, AZ, but held more confidence in
elections administration in their local area. This complements previous
research that demonstrates a similar bias favorable to one's local area
(\citeproc{ref-stewart2022}{Stewart 2022}).

@@

\begin{tabular}[t]{llllllllll}
\toprule
Place & $\bar{x}_{diff}$ & $\bar{x}_{treat}$ & $\bar{x}_{control}$ & $treat_{n}$ & $control_{n}$ & t & p & data & CI\\
\midrule
Maricopa County, AZ & 0.20 & 0.93 & 0.72 & 650.00 & 637.00 & 2.96 & 0.003** & 1283.73 & {}[ 0.07, 0.33]\\
Local Area & 0.06 & 1.18 & 1.12 & 650.00 & 637.00 & 0.87 & 0.383 & 1282.95 & {}[-0.08, 0.20]\\
\bottomrule
\multicolumn{10}{l}{\rule{0pt}{1em}\textit{Note: }}\\
\multicolumn{10}{l}{\rule{0pt}{1em}Welch Two Sample t-tests of difference between Confidence in Elections by Treatment Condition.}\\
\end{tabular}
\endgroup{}

@@

Among those who refuted the 2020 election results, trust in Maricopa
County, AZ elections was lower for each item compared to items
pertaining to elections within one's local area, regardless of treatment
condition. However, notably, trust that election technology would be
secure received the lowest confidence endorsement overall. There was
also a notable lack of trust that the voting process would be fair
relative to the other items.\footnote{What is interesting to note here
  is that the first three items regard the more human element of the
  election administration process, e.g., ensuring accuracy of counts,
  competence, and commitment of election workers. Although a neat line
  can't really be drawn between the first three and latter two items
  presented here, it is interesting to see a slight divergence from the
  usual high confidence placed in local area elections.} Now the
comparative difference in responses between those in the treatment group
compared to those in the control group appears substantial for items
pertaining to elections in Maricopa County, AZ.

@@

Concerning distrust among those who refuted the 2020 election results,
responses items assessing one's expectation that election fraud would
occur reveal the same pattern. Although there is not as stark a
disparity between responses of those in the treatment and those in the
control, among those in the treatment group, there does appear to be
less expectation that election fraud would occur in Maricopa County, AZ.
Overall, review of responses to trust (Figure~\ref{fig-patch13}) and
distrust survey items (Figure~\ref{fig-patch14}) does lend credence to
the impact of the treatment.

@@

@@

>>>>

% ---------------------------
% END CUSTOM PREAMBLE STUFF
% ---------------------------
\usepackage{float}
\usepackage{tabularray}
\usepackage[normalem]{ulem}
\usepackage{graphicx}
\usepackage{rotating}
\UseTblrLibrary{booktabs}
\UseTblrLibrary{siunitx}
\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{dcolumn}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Inspiring Confidence and Easing Insecurity in the Pre-election period: The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={RoyalBlue},
  pdfcreator={LaTeX via pandoc}}

\title{Inspiring Confidence and Easing Insecurity in the Pre-election
period: The Effect of Recruiting Military Veterans on Confidence in 2024
US Elections}
% subtitles do not seem to work with article class?
%%\subtitle{}

@@

% add the short title to the fancy header
\fancyhead[R]{Effect of Recruiting Military Veterans on Confidence in
2024 US Elections}
\fancyhead[L]{Isaiah}

@@

\setstretch{1.25}
\newpage{}

@@

Election administration officials make efforts to sustain public trust
and confidence in the fairness and accuracy of elections, and attempt to
boost such confidence where it may be deprived. Concerns for safety have
developed among election staff and voters in more recent elections.
Regular measures are taken to enhance the \emph{trustworthiness} of the
electoral process through practices meant to improve the conduct,
transparency, or overall administration of elections in the United
States.

Although election officials undertake great efforts to enhance the
\emph{trustworthiness} of election administration, public
\emph{trust}\footnote{I adopt a distinction made between
  \emph{trustworthiness} and \emph{trust} in elections
  (\citeproc{ref-stewart2022}{Stewart 2022}). The ``\emph{worthiness}''
  of one's trust in the conduct and administration of elections is based
  on the extent that outcomes of an election reasonably follow the rules
  prescribed and can be adjudicated as such. The public's \emph{trust}
  in elections, however, is amendable to an indefinite number of factors
  that may be unrelated to the formal structure or procedures of
  election administration. I can recognize that my car is trustworthy
  prior to ever driving it because it is structurally sound; I have
  every reason to deem it \emph{worthy of my trust}. However, I don't
  trust my car because I am sure it is haunted.} in elections is a
psychological construct influenced by many things outside of election
official control such as partisanship or elite rhetoric
(\citeproc{ref-hooghe2018}{Hooghe 2018};
\citeproc{ref-sances2015}{Sances and Stewart 2015}). Moreover, a
person's evaluation of the election in hindsight is often influenced by
the election outcome itself (\citeproc{ref-daniller2019}{Daniller and
Mutz 2019}; \citeproc{ref-stewart2022}{Stewart 2022}). Thus, measures
taken by election officials can be undermined, trivialized, or made
irrelevant depending on how one feels after the election results have
come out.

Such volatile attitudes and evaluations post-election can leave a
lasting impression that election officials must contend with upon the
next election cycle (\citeproc{ref-bowler2024}{Bowler and Donovan 2024};
\citeproc{ref-levendusky2024}{Levendusky et al. 2024}). For instance, we
have witnessed many people's outright refusal to accept the 2020 U.S.
election results as legitimate despite consistent review of the evidence
confirming the results as fair and accurate. Such a case demonstrates
that public trust in elections is, at best, only partial to
trustworthiness of election administration in the United States.

@@

In addition, prior research has shown that confidence that one's own
vote would be counted as intended has been much stronger than confidence
that votes nationwide would be treated likewise
(\citeproc{ref-sances2015}{Sances and Stewart 2015};
\citeproc{ref-stewart2022}{Stewart 2022}). In other words, the public is
more trusting of elections closer to home. Taking into account that
confidence is lower for elections conducted outside of one's local area,
then information that veterans are being recruited to help conduct
elections in locations beyond one's locale may close this gap between
confidence in elections within one's local area and confidence in
elections elsewhere. Although it may be difficult to discern whether
recruiting veteran service members would improve confidence in elections
near one's local area in light of such hometown favoritism, a positive
effect on confidence in elections held beyond one's local area---where
confidence is already expectedly lower---can help discern the impact of
the group alone on public opinion. Therefore, any observed disparity of
confidence in elections between those within one's local area and those
beyond will be smaller among those who are presented with information
about election official recruitment efforts of veteran service members.

Since the results and events following the 2020 election loomed large in
anticipation of the U.S. 2024 general elections, it is reasonable to
take into account beliefs about legitimacy of the 2020 election. This
group who denies the legitimacy of the 2020 election results were
possibly the most apprehensive (distrusting) of the 2024 election.
However, given that veteran service members are held in high regard,
generally speaking, announcing that veterans are being actively
recruited to work as election staff and volunteers is expected to
positively influence confidence in elections among those who remained
firm in the belief that the 2020 election results were illegitimate.
That is to say, despite their belief about the illegitimacy of the 2020
election results, I expect that informing them about veteran recruitment
efforts to improve their confidence in elections. I take into account
the relationship between partisanship and 2020 election legitimacy
beliefs and include partisanship as a control in the
analysis\footnote{To be clear, I conducted an auxiliary regression
  analysis to examine the relationship between legitimacy beliefs and
  party ID. Results revealed a statistically significant relationship
  between the two variables and thus omitting either would result in
  biased estimates.}.

Finally, I expect that concerns about potential for violence will
lessen, and confidence in voter safety improve, when presented with
information that veterans are being actively recruited to work and
volunteer in election offices.

@@

Survey participants read either a treatment or control vignette, which
resembled a news article about efforts in Maricopa County, AZ to recruit
election staff and volunteers for the 2024 general election\footnote{The
  article vignettes were written and designed by Michael Hanmer Ph.D and
  myself with input from We the Veterans, a non-profit organization}.
The treatment vignette referred to a program designed to recruit
veterans and their family members and describes an interviewee ``Jordan
Braxton'' as an Army veteran. The control vignette simply omitted any
mention of veterans and their family members, and didn't describe
``Jordan Braxton'' as an Army veteran. Beyond those small differences
and the headlines, the article vignettes are identical. Therefore,
effects can be attributed to the information about veterans in the
treatment vignette\footnote{Complete text of treatment and control
  vignettes are included in Appendix A}.

@@

\centering
\begin{tblr}[         %% tabularray outer open
]                     %% tabularray outer close
{                     %% tabularray inner open
width={1\linewidth},
colspec={X[0.5]X[0.5]},
column{1,2}={}{font=\fontsize{0.8em}{1.1em}\selectfont,},
}                     %% tabularray inner close
\toprule
Trust & Distrust \\ \midrule %% TinyTableHeader
How confident are you that votes in Maricopa County, AZ will be counted as voters intend in the elections this November? & There will be voter fraud, that is, people who are not eligible to vote will vote, or vote more than once \\
How confident are you that election officials, their staff, and volunteers in Maricopa County, AZ will do a good job conducting the elections this November? & Many votes will not actually be counted \\
How commited do you think election staff and volunteers in Maricopa County, AZ will be to making sure the elections are fair and accurate? & Many people will show up to vote and be told they are not eligible \\
How confident are you that the voting process will be fair in Maricopa County, AZ? & A foreign country will tamper with the votes cast in this area to change the results \\
How confident are you that election systems in Maricopa County, AZ will be secure from hacking and other technological threats? & Election officials in Maricopa County, Arizona will try to discourage some people from voting \\
\bottomrule
\end{tblr}

@@

Upon construction of the two summated scales of \emph{trust} and
\emph{distrust}, internal consistency reliability coefficient \(\alpha\)
were estimated for both and found to be adequate (\(\alpha\)
\textgreater{} 0.8); the trust in elections scale \(\alpha\) = 0.93 and
the distrust scale \(\alpha\) = 0.86. In other words, the items in each
of the scales were intercorrelated well enough for each to compose a
single scale presuming a unidimensional construct for each\footnote{See
  Appendix C for Inter-item correlation analysis.}.

@@

In light of this, I reverse coded items on the distrust scale and
combined responses to the trust in elections scale. That is, item
responses on the trust scale ranged from 0 (``Not confident at all'') to
3 (``Very confident), whereas responses on the distrust scale ranged
from 0 (''Not likely at all'') to -3 (``Very likely''). Combining
responses in this way resulted in an overall measure of \emph{confidence
in elections} to serve as the dependent variable\footnote{The
  methodological concerns between the use of (unit weighted) sum or mean
  scores compared to the use of estimated factor scores for Likert
  scales composed of multiple Likert items are legion. The use of either
  computed sum scores or estimated factor scores as a general practice
  is continuously debated (\citeproc{ref-mcneish2023}{McNeish 2023};
  \citeproc{ref-mcneish2020}{McNeish and Wolf 2020};
  \citeproc{ref-widaman2022}{Widaman and Revelle 2022},
  \citeproc{ref-widaman2024}{2024}), yet there is no general consensus
  on the most appropriate method generally. The use of either appears to
  largely depend on the researcher's particular research objective,
  theoretical suppositions, and necessity. Another common concern is the
  inappropriate treatment of ordinal variables as interval which levy
  strong assumptions, e.g., each item contributes equally to the
  variable being measured and that ordinal response categories are
  equidistant as though on an interval scale
  (\citeproc{ref-wang2013}{Wang et al. 2013}; for contrast, see
  \citeproc{ref-robitzsch2020}{Robitzsch 2020}).}. On this scale,
positive values reflect higher confidence in elections (e.g., elections
will be fair and accurate), negative values reflect greater distrust
(i.e., expectation that election fraud will occur), and values closer to
zero correspond to an individual's degree of \emph{insecurity}.

@@

Once trust and distrust were composed into unified variable of
confidence in elections, I examined distribution of the dependent
variable followed by analysis of the effect of the treatment by
comparison of mean differences between experiment conditions. To retain
a meaningful zero, I rescaled the sum scores to range from -3 to
3\footnote{This range is mostly arbitrary, as a range from -1 to 1 works
  much the same. When relying on simple sum total score, the scale
  ranges from -15 to 15. So a single point increase from 0 to 1 may
  reflect a combination of a single ``Very confident'' response on a
  trust item compensated by a single ``Somewhat likely'' response on a
  distrust item (i.e., 3-2 = 1), or some other equivalent combination.
  Although the use of mean scores would place scores back onto the
  response scale metric (e.g., from 0 to 3, reflective of ``not at all''
  to ``very''), this results in a unipolar scale (from 0 to 3). The
  bipolar scale resulting from the composition of positive trust and
  negative distrust engenders meaning to zero and negative values. In
  other words, negative values on the scale for confidence in elections
  hold substantive meaning (e.g., distrust, absent trust), which makes
  transforming scores to fit a unipolar scale inappropriate.} and then
superimposed two histograms of confidence scores: one derived from AZ
items and another for scores derived from local area items
(Figure~\ref{fig-dist}). Doing this is illustrative of the relative
distribution of confidence that differed depending on the location of
elections referred to in the survey items (e.g., Maricopa County, AZ).
As expected, the distribution of confidence differed depending on
whether survey items pertained to elections in Maricopa County, AZ, or
elections within one's local area.

@@

Overall, the sample expressed more trust in elections relative to
distrust seeing as how the distribution of confidence is largely
positive. However, there's a clear difference between confidence in
Maricopa county, AZ elections and elections within one's local area. A
paired t-test confirms that this difference was statistically
significant, though somewhat small (mean difference \(= 0.325\),
\(95\%\) CI \([0.27, 0.38]\), \(t(1286) = 11.61\), \(p < .001\)).
Nonetheless, respondents appeared less confident regarding elections in
Maricopa County, AZ, but held more confidence in elections
administration in their local area. This complements previous research
that demonstrates a similar bias favorable to one's local area
(\citeproc{ref-stewart2022}{Stewart 2022}).

@@

\centering
\begin{tblr}[         %% tabularray outer open
]                     %% tabularray outer close
{                     %% tabularray inner open
colspec={Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]Q[]},
column{1,2,3,4,5,6,7,8,9,10}={}{font=\fontsize{0.75em}{1.05em}\selectfont,},
}                     %% tabularray inner close
\toprule
Place & $\bar{x}_{diff}$ & $\bar{x}_{treat}$ & $\bar{x}_{control}$ & $n_{treat}$ & $n_{control}$ & t & p & df & CI \\ \midrule %% TinyTableHeader
Maricopa County, AZ & $0.20$ & $0.93$ & $0.72$ & $650.00$ & $637.00$ & 2.96 & 0.003** & 1283.73 & [ 0.07, 0.33] \\
Local Area & $0.06$ & $1.18$ & $1.12$ & $650.00$ & $637.00$ & 0.87 & 0.383 & 1282.95 & [-0.08, 0.20] \\
\bottomrule
\end{tblr}

@@

Among those who refuted the 2020 election results, trust in Maricopa
County, AZ elections was lower for each item compared to items
pertaining to elections within one's local area, regardless of treatment
condition. However, notably, trust that election technology would be
secure received the lowest confidence endorsement overall. There was
also a notable lack of trust that the voting process would be fair
relative to the other items\footnote{What is interesting to note here is
  that the first three items regard the more human element of the
  election administration process, e.g., ensuring accuracy of counts,
  competence, and commitment of election workers. Although a neat line
  can't really be drawn between the first three and latter two items
  presented here, it is interesting to see a slight divergence from the
  usual high confidence placed in local area elections.}. Now the
comparative difference in responses between those in the treatment group
compared to those in the control group appears substantial for items
pertaining to elections in Maricopa County, AZ.

@@

Concerning distrust among those who refuted the 2020 election results,
responses to items assessing one's expectation that election fraud would
occur reveal the same pattern. Although there is not as stark a
disparity between responses of those in the treatment and those in the
control, among those in the treatment group, there does appear to be
less expectation that election fraud would occur in Maricopa County, AZ.
Overall, review of responses to trust (Figure~\ref{fig-patch13}) and
distrust survey items (Figure~\ref{fig-patch14}) does lend credence to
the impact of the treatment.

@@

\caption{\label{tbl-legit}Confidence in Elections by Treatment and
Legitimacy Beliefs}

\begin{tabular}{l c c}
\hline
 & Confidence in AZ Elections & w/Interaction \\
\hline
(Intercept)                & $0.03 \; (0.08)$        & $0.07 \; (0.08)$        \\
Treatment                  & $0.18 \; (0.05)^{***}$  & $0.09 \; (0.06)$        \\
Not Legitimate             & $-0.72 \; (0.07)^{***}$ & $-0.86 \; (0.09)^{***}$ \\
Republican                 & $0.11 \; (0.08)$        & $0.12 \; (0.08)$        \\
Democrat                   & $0.17 \; (0.09)$        & $0.17 \; (0.09)^{*}$    \\
Not Legitimate * Treatment &                         & $0.27 \; (0.11)^{*}$    \\
\hline
R$^2$                      & $0.14$                  & $0.14$                  \\
Adj. R$^2$                 & $0.13$                  & $0.14$                  \\
Num. obs.                  & $1263$                  & $1263$                  \\
\hline
\multicolumn{3}{l}{\scriptsize{$^{***}p<0.001$; $^{**}p<0.01$; $^{*}p<0.05$}}
\end{tabular}

@@

\subsubsection{Concerns for Violence and Voter
Safety}\label{concerns-for-violence-and-voter-safety}
